---
title: "Trial_GPS_AccessibilityPrep"
author: "Brandon Mayer"
date: "2024-06-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
```

Load in point data and shapefile for clipping
```{r}
# Example point data


# Example polygon (assuming a simple square for the purpose of this example)
poly_data <- rbind(c(10, 10), c(90, 10), c(90, 90), c(10, 90), c(10, 10))
polygon <- vect(poly_data, type="polygons", crs="+proj=longlat +datum=WGS84")
```

Clip points by the polygon:
Use the crop function to clip the points to the area within the polygon.
```{r}
# Clip points within the polygon
clipped_points <- crop(points, polygon)
```

Create an empty raster and rasterize the points:
Define a raster based on the extent of the polygon or a specified area and rasterize the points.
```{r}
# Define the extent from the polygon or manually
raster_extent <- ext(polygon)  # Using the polygon's extent

# Create an empty raster
r <- rast(raster_extent, nrow=10, ncol=10, vals=0)

# Rasterize points, counting the number in each cell
count_raster <- rasterize(clipped_points, r, fun=sum)
```

Calculate the percentage of total points per cell:
Calculate the total number of points and then convert the counts to percentages.
```{r}
# Calculate total number of clipped points
total_points <- nrow(clipped_points)

# Compute the percentage raster
percentage_raster <- count_raster
values(percentage_raster) <- (values(count_raster) / total_points) * 100
```

Visualization:
Plot the resulting raster with the polygon overlay to visualize the distribution of point percentages.
```{r}
# Plot the percentage raster
plot(percentage_raster, main="Percentage of Total Points per Cell", col=rainbow(100, alpha=0.5))
lines(polygon, col="black", lwd=2)  # Add polygon outline for clarity
```

Notes:
Handling NA values: If there are cells in the raster with no points, ensure these are handled properly (e.g., set to 0 or kept as NA depending on your analysis needs).
Spatial extent and resolution: Adjust these parameters based on the spatial scale and detail you need for your analysis.
Visualization enhancements: Adjust the color palette and other visualization parameters to best represent your data.
Using terra for these tasks is efficient for handling large spatial data, and the workflow integrates spatial operations like clipping and rasterization seamlessly.
